---
title: "assignment7"
format: gfm
editor: visual
---

```{r}
library(tidyverse)
library(knitr)
library(dslabs)
```

## **Excercise: 2016 election result and polling**

For this exercise, we will explore the result of the 2016 US presidential election as well as the polling data. We will use the following three datasets in the `dslabs` package, and use `join` function to connect them together. As a reminder, you can use `?` to learn more about these datasets.

-   `results_us_election_2016`: Election results (popular vote) and electoral college votes from the 2016 presidential election.

-   `polls_us_election_2016`: Poll results from the 2016 presidential elections.

-   `murders`: Gun murder data from FBI reports. It also contains the population of each state.

We will also use [this dataset](https://raw.githubusercontent.com/kshaffer/election2016/master/2016ElectionResultsByState.csv) to get the exact numbers of votes for question 3.

### **Question 1. What is the relationship between the population size and the number of electoral votes each state has?**

**1a.** Use a `join` function to combine the `murders` dataset, which contains information on population size, and the `results_us_election_2016` dataset, which contains information on the number of electoral votes. Name this new dataset `q_1a`, and show its first 6 rows.

```{r}
q_1a <- left_join(murders,results_us_election_2016) 

q_1a |> head(6) |> kable()
```

**1b.** Add a new variable in the `q_1a` dataset to indicate which candidate won in each state, and remove the columns `abb`, `region`, and `total`. Name this new dataset `q_1b`, and show its first 6 rows.

```{r}
q_1b <- q_1a |> 
select(-abb) |> 
mutate(winner=ifelse(clinton > trump, "clinton", "trump"))

q_1b |> head(6) |> kable()
```

**1c.** Using the `q_1b` dataset, plot the relationship between population size and number of electoral votes. Use color to indicate who won the state. Fit a straight line to the data, set its color to black, size to 0.1, and turn off its confidence interval.

```{r}
q_1b |> ggplot(mapping=aes(x=population, y=electoral_votes)) + 
geom_line() + geom_point(mapping=aes(color=winner))
```

### **Question 2. Would the election result be any different if the number of electoral votes is exactly proportional to a state's population size?**

**2a.** First, convert the `q_1b` dataset to longer format such that the `population` and `electoral_votes` columns are turned into rows as shown below. Name this new dataset `q_2a`, and show its first 6 rows.

```{r}
q_2a <- q_1b |> 
pivot_longer(c(electoral_votes,population),names_to="metric",values_to="value") 

q_2a |> head(6) |> kable()
```

**2b.** Then, sum up the number of electoral votes and population size across all states for each candidate. Name this new dataset `q_2b`, and print it as shown below.

```{r}
q_2b <- 
q_2a |> 
select(metric,winner,value) |> 
group_by(metric,winner) |> 
summarize(value=sum(value))

kable(q_2b)
```

**2c. Use the q_2b dataset to contruct a bar plot to show the final electoral vote share under the scenarios of 1) each state has the number of electoral votes that it currently has, and 2) each state has the number of electoral votes that is exactly proportional to its population size. Here, assume that for each state, the winner will take all its electoral votes.**

```{r}
q_2b |> 
ggplot(mapping = aes(x=metric,y=value,color=winner))+geom_col(position = "fill")
```

\*\*3a. First, from this dataset on GitHub, calculate the number of popular votes each candidate received as shown below. Name this new dataset q_3a, and print it.

Note: Vote counts are listed for several other candidates. Please combine the votes for all candidates other than Clinton and Trump into a single others category (as shown in the table below)

Hint: pivot_longer() may be useful in here.\*\*

```{r}
votes <-read.csv ("https://raw.githubusercontent.com/kshaffer/election2016/master/2016ElectionResultsByState.csv")

q_3a <- 
votes |> 
summarize(clinton=sum(clintonVotes+clintonElectors),
trump=sum(trumpVotes+trumpElectors),
others=sum(steinVotes+mcmullinVotes+othersVotes)) |> 
mutate(metric = "popular_votes") |>
relocate(metric) |> 
pivot_longer(c(clinton,trump,others),names_to ="winner",values_to="value") 

head(q_3a) |> kable() 

```

**3b. Combine the `q_2b` dataset with the `q_3a` dataset. Call this new dataset `q_3b`, and print it as shown below.**

```{r}
q_3b <- 
bind_rows(q_2b,q_3a) 

kable(q_3b)
```

**3c. Lastly, use the `q_3b` dataset to contruct a bar plot to show the final vote share under the scenarios of 1) each state has the number of electoral votes that it currently has, 2) each state has the number of electoral votes that is exactly proportional to its population size, and 3) the election result is determined by the popular vote.**

```{r}
q_3b |> 
ggplot(mapping = aes(x=metric,y=value,color=winner))+geom_col(position = "fill")
```

**4a. The polling data is stored in the data frame `polls_us_election_2016.` For the sake of simplicity, we will only look at the data from a single poll for each state. Subset the polling data to include only the results from the pollster Ipsos. Exclude national polls, and for each state, select the polling result with the enddate closest to the election day (i.e. those with the lastest end date). Keep only the columns state, adjpoll_clinton, and adjpoll_trump. Save this new dataset as q_4a, and show its first 6 rows.**


*Note: You should have 47 rows in q_4a because only 47 states were polled at least once by Ipsos. You don’t need to worry about the 3 missing states and DC.

Hint: `group_by()` and `slice_max()` can be useful for this question. Check out the help file for `slice_max()` for more info*

```{r}
q_4a <- polls_us_election_2016 |>  
filter(pollster == "Ipsos",state != "U.S.") |> 
group_by(state) |> 
slice_max(startdate) |> 
select(state,adjpoll_clinton,adjpoll_trump)

head(q_4a) |> kable()
```

